\Large\textbf{Preparation GUIDE for Initial-Submission Manuscript}\\~\\ \large\textbf{ Preparation of Papers in Two-Column Format\\ for the ASP-DAC 2014 (\LaTeX2e version)}

### 2013-07-01 background.tex
My changes
Included all changes by Prof. Salcic

\begin{figure*}[bth]
\centering
\begin{SubFloat}{\label{fig:2a}Simple SystemJ program}%...verbatim subfigure
\begin{minipage}[b]{0.3\linewidth}% a minipage to control the width...
\begin{verbatim}
while(true){
 abort(A){
  while(true)
   pause;
  emit O;
  // do java computation
}
\end{verbatim}%
\end{minipage}%
\end{SubFloat}
\hspace{1.5cm}%
\begin{SubFloat}[Black box]{\label{fig:2b}Logical time in SystemJ}%
\includegraphics[scale=0.4]{moc}
\end{SubFloat}%
\begin{SubFloat}{\label{fig:2c}Mapping to physical time}%
% \includegraphics[scale=0.4]{phy.pdf_t}
\scalebox{0.4}{\input{phy.pdf_t}}
\end{SubFloat}%
\caption{Simple SystemJ example and corresponding MoC}
\label{fig:2}
\end{figure*}

SystemJ consists of entities called clock-domains (CD)s, each running at
their individual pace, interacting with the environment via signals and
with each other via channels and CSP~\cite{pwel05} style
rendezvous. Each CD adheres to the \textit{perfect synchrony}
hypothesis, i.e., all statements execute instantaneously in zero
time. Only the \texttt{pause} statement consumes time, just like in
Esterel~\cite{gber931}. Consider the SystemJ program in
Figure~\ref{fig:2a}, it is waiting for an input signal producing logical
ticks (see Figure~\ref{fig:2b}). Once \texttt{A} is received at tick 4,
output \texttt{O} is emitted to the environment instantaneously.

@@

The perfect synchrony hypothesis is ideal for programming the
synchronous sub-set of SystemJ. But, the real-world is not that
forgiving, execution of every \textit{micro-step} in the logical zero
time, requires $\delta$ physical time. Time for a complete reaction to
one or more input signal can thus be summarized as $\Delta = \Sigma
(\delta)$. We call $\Delta$ the reaction time. Obviously, depending upon
the amount of computation required, $\Delta$ can vary, hence, in order
to satisfy the implicit restriction imposed by the synchrony hypothesis,
no input event can be missed, one needs to calculate the \textit{Worst
  Case Reaction Time} (WCRT) and the resultant WCRT needs to be smaller
than the speed of the fastest input event, else the synchrony hypothesis
is violated. A substantial amount of research for calculating the WCRT
of synchronous programs~\cite{proop10,boldt07,wilhelm08} exists. We
refer the reader to any of these research articles, since WCRT analysis
is not the focus of this paper. The opposite of the WCRT is the
\textit{Best Case Reaction Time} (BCRT), we denote the WCRT and BCRT in
Figure~\ref{fig:2c} using \texttt{W} and \texttt{B},
respectively. Figure~\ref{fig:2c} shows a continuously running physical
clock (analogous to a clock in hardware), the numerical annotations on
the rising edge of the clock mark the logical ticks from
Figure~\ref{fig:2b}. The subscripts \texttt{s} and \texttt{e} for the
numbers show the start and end of the logical ticks, respectively. The
end of a logical tick and the start of the next logical tick happen
together. Figure~\ref{fig:2c} clearly shows the mapping of the logical
time to the physical time. As we can see due to Java computations
(Figure~\ref{fig:2a}) the physical time between logical ticks 4 and 5 is
2 clock cycles, while for the rest of the ticks it is just 1 clock
cycle. This elasticity is inherent (and elegant) to both Esterel and
SystemJ.

>>>>

SystemJ consists of entities called clock-domains (CD)s, each running at
their individual pace, interacting with the environment via signals and
with each other via channels and CSP~\cite{choa85} style
rendezvous. Each CD has its own notion of logical time. CDs adheres to
the \textit{perfect synchrony} hypothesis, i.e., all statements execute
instantaneously in zero time. Only the \texttt{pause} statement consumes
time, just like in Esterel~\cite{gber931}. Consider the SystemJ program
in Figure~\ref{fig:2a}, it is waiting for an input signal producing
logical ticks (see Figure~\ref{fig:2c}). Once \texttt{A} is received at
tick 1 or 4, output \texttt{O} is emitted to the environment
instantaneously.

@@

\begin{figure}[t!]
% \centering
\begin{SubFloat}{\label{fig:2a}Simple SystemJ program}%...verbatim subfigure
\begin{minipage}[b]{0.3\linewidth}% a minipage to control the width...
  \begin{scriptsize}
\begin{verbatim}
{
int counter = 0;
while(true){
 abort(A){
  while(true) pause; 
  emit O;
  if(counter != 0) { // do java computation leading to WCRT}
  ++counter;
  } // abort end
 } // loop end
} // SystemJ CD end
\end{verbatim}%
  \end{scriptsize}
\end{minipage}%
\end{SubFloat}
\hspace{1.5cm}%
% \begin{SubFloat}[Black box]{\label{fig:2b}Logical time in SystemJ}%
% \includegraphics[scale=0.4]{moc}
% \end{SubFloat}%
% \vspace{-10pt}

\begin{SubFloat}{\label{fig:2c}Mapping to physical time}%
% \includegraphics[scale=0.4]{phy.pdf_t}
\scalebox{0.68}{\input{phy.pdf_t}}
\end{SubFloat}%
\caption{Simple SystemJ example and corresponding MoC}
\label{fig:2}
\end{figure}

The perfect synchrony hypothesis is ideal for programming the
synchronous sub-set of SystemJ. Unfortunately, every \textit{micro-step}
requires real-time to compute. Formally, execution of every micro-step
in the logical zero time, requires $\delta$ physical time. Time for a
logical tick can thus be summarized as $\Delta = \mathcal{F} (\delta)$,
where $\mathcal{F}$ is some function dependent upon the schedule. We
call $\Delta$ the reaction time. Obviously, depending upon the amount of
computation required and schedule, $\Delta$ can vary, hence, in order to
satisfy the implicit restriction imposed by the synchrony hypothesis --
no input event \textit{from the environment} can be missed -- one needs
to calculate the \textit{Worst Case Reaction Time} (WCRT) and the
resultant WCRT needs to be smaller than the time between any two
consecutive events on any input, else the synchrony hypothesis is
violated. Techniques for calculating the WCRT of synchronous
programs~\cite{boldt07} exist and this analysis is not the focus of this
paper. The opposite of the WCRT is the \textit{Best Case Reaction Time}
(BCRT), we denote the WCRT and BCRT in Figure~\ref{fig:2c} using
\texttt{W} and \texttt{B}, respectively. Figure~\ref{fig:2c} shows a
continuously running physical clock (analogous to a clock in digital
hardware), the numerical annotations on the rising edge of the clock
mark the logical ticks from Figure~\ref{fig:2c}. The subscripts
\texttt{s} and \texttt{e} for the numbers show the start and end of the
logical ticks, respectively. The end of a logical tick and the start of
the next logical tick happen together. Figure~\ref{fig:2c} shows the
mapping of the logical time to the physical time. For example, the
logical tick \texttt{1} starts at the first rising edge and completes at
the second rising edge of the physical clock. Where as logical tick
\texttt{4} starts at the $4^{th}$ rising edge, but finishes at the
$6^{th}$ rising edge of the physical clock. This difference can be
accounted for by the java computations required when signal \texttt{A}
is input by the environment (see Figure~\ref{fig:2a}). This elasticity
is inherent (and elegant) to both Esterel and SystemJ.

### 2013-07-01 conclusion.tex
first commit
Included all changes by Prof. Salcic

\section{Background}
\label{sec:background}

>>>>

\section{Conclusion}
\label{sec:concl-future-work}

In this paper we have described a novel way to introduce real-time in
\textit{Globally Asynchronous Locally Synchronous} (GALS) languages, in
particular SystemJ, and their subset the synchronous languages like
Esterel. The fundamental idea is to convert real-time delay into logical
delays (\texttt{pause} construct) that interact well with rest of the
constructs in these languages, especially, preemption and channel
communication. We do \textit{not} need or use external timers to
introduce these real-time delays, thereby resolving the problems of
interaction between external timers and preemption. Moreover, we also do
not require specific timer resolutions, thereby making our solution
elegant (in the spirit of GALS/synchronous programming) and flexible. We
rewrite the delay construct into reactive kernel constructs, calculate
the \textit{Worst} and \textit{Base} case reaction times and then
determine \textit{delay} \texttt{d}, number of logical ticks, which is
invariant to the \textit{elasticity} of logical time. To our knowledge
we are the first to introduce non-deterministic delays in such languages
and show their usefulness in designing real-world systems.

### 2013-07-01 delays.tex
Included all changes by Prof. Salcic

\section{Introducing real time in SystemJ}
\label{sec:intr-real-time}

We introduce a single \textit{derived} construct called
\mbox{\texttt{delay [M..N]}} in the SystemJ language for real time
control. The resolution of the delay statement is of secondary concern
and is dependent upon the execution platform, without loss of generality
through out this paper we consider the resolution to be always in milli
seconds.

@@

Given a SystemJ program: \texttt{delay [M..N]; p}, where $M \in
\mathbb{Q}^{>0}$ and $N \in \mathbb{Q}^{>0}$, statement $p$ is executed
after real-time $\tau$, such that, $M \leq \tau \leq N$.

We also introduce two variants of the derived statement \texttt{delay}.
\begin{enumerate*}
\item Given a SystemJ program \texttt{delay M; p}, where $M \in
  \mathbb{Q}^{>0}$, statement $p$ is executed after real-time $\tau$,
  such that, $M \leq \tau \leq \infty$. It is important to note that the
  lower bound of the delay construct $M$ is \textit{not} an exact delay,
  but rather the control is allowed to proceed to the next statement
  anytime after the delay of time $M$.
\item Give a SystemJ program \texttt{delay [M..M]; p}, where $M \in
  \mathbb{Q}^{>0}$, statement $p$ is executed after real-time $\tau$,
  such that, $M \leq \tau \leq M$. In this variant, the delay is exact.
\end{enumerate*}

The aforementioned variants are specializations of the general case
\texttt{delay [M..N]}.

@@

\begin{figure}[tb]
    \begin{minipage}{\textwidth}
      \begin{scriptsize}
\begin{verbatim}
trap(T){
 int x = 0;
 while(true){
  x = x + 1;
  if(x == d) exit (T); //wait for "d" ticks
  pause;
 }
}
\end{verbatim}
      \end{scriptsize}
    \end{minipage}
    \caption{The rewrite of \texttt{delay} construct}
    \label{fig:3}
\end{figure}

@@

The computation of \texttt{d} is shown in Algorithm~\ref{alg:1}. This
algorithm is carried out for each CD (in SystemJ) or a synchronous
program individually. \textit{The fundamental observation is -- the
  reaction time for each logical tick is elastic -- varying only between
  the BCRT and the WCRT, thus any logical number of ticks \texttt{d}
  that map to the required real-time \texttt{delay} should be chosen in
  such a way that they are invariant to this elasticity.}

Our algorithm takes as input: WCRT, BCRT, and the lower and upper bounds
$M$ and $N$, respectively of the \texttt{delay} construct. We first
divide $M$ and $N$ with the BCRT and WCRT, respectively. This division
gives us the number of individual logical ticks required to delay the CD
(or synchronous program) by the real-time specification. We always
$ceil$ when dividing $M$ and $floor$ when dividing $N$ to make sure that
the resultant values are integers (in domain $\mathbb{N}^{>0}$) and
these functions guarantee that the resultant logical ticks result in
real-time delays between the required range $M-N$. Next, a linear
function $F$ maps these calculated values to a set of equidistant
integer points (values) separated by a unit value -- these points
represent all the logical ticks running at the WCRT and the BCRT,
respectively that satisfy the real-time delay requirements. The
intersection of these two sets gives all the logical ticks that satisfy
the real-time requirements invariant of the reaction time and its
elasticity.

Let us revisit our motivating example to elucidate the algorithm. From
Figure~\ref{fig:1} we know that $M$ is 50ms and $N$ is 100ms,
respectively. Let the WCRT and BCRT be: 13.332ms and 10ms,
respectively. Thus, the algorithm proceeds as follows:

\begin{enumerate*}
\item $l_1 \leftarrow \lceil 50/13.332 \rceil$ and $u_1 \leftarrow
  \lfloor 100/13.332 \rfloor$. $l_1 = 4$ and $u_1 = 7$. We first
  calculate the logical ticks that are always running at the WCRT
  satisfy the required real-time delay.
\item $l_2 \leftarrow \lceil 50/10 \rceil$ and $u_2 \leftarrow
  \lfloor 100/10 \rfloor$. $l_1 = 5$ and $u_1 = 10$. We do the same for
  the BCRT case.
\item $S_1 = \{4,5,6,7\}$ and $S_2 =\{5,6,7,8,9,10\}$. We then map the
  resultant bounds to linear points. Sets $S_1$ and $S_2$ represent
  logical ticks that running at the WCRT and BCRT, respectively always
  satisfy the required real-time constraints.
\item $D = S_1 \cap S_2$, $D = \{5,6,7\}$. Finally, the intersection of
  the two sets gives the set $D$ from which we can pick up any value for
  \texttt{d}.
\end{enumerate*}

The resultant value for \texttt{d} gives the number of logical ticks,
which can run at any physical clock-speed, bounded by the BCRT and the
WCRT of the program and still result in the required real-time delay. We
think this is an elegant solution, because the technique provides
hard-real time guarantee while preserving the essence of synchronous and
GALS programming prescribed by SystemJ and Esterel style
languages. Moreover, this technique considers non maximal-parallelism,
i.e., the delays in logical ticks is calculated after scheduling has
been performed. To our knowledge we are the first to do so.

@@

A number of programs require non-deterministic timing constructs. One of
them is our motivating example -- the human response time
system. Another is a printer-spooler example borrowed from timed
CSP~\cite{Schneider:1999:CRT:555233}. The spooler and printer need to
synchronize using channels. The printer might be unable to print
depending upon the paper tray, similarly, spooler might take sometime to
send the job depending upon its size. Such real-time constraints can be
modeled in SystemJ as below:

\begin{scriptsize}

\begin{verbatim}
{await (job); delay([2..10]ms);  //SPOOLER CD
 send mid(job) delay (1ms);}
><
{receive mid; delay([1..30]ms); //PRINTER CD
 emit print(#mid);}
\end{verbatim}
\end{scriptsize}

@@

There are many instances when one would like to wait on an input from
environment for only a specified amount of time. This can be programmed
as:
    \begin{scriptsize}
\begin{verbatim}
// timeout after 1ms.
trap(T){
 {await(A);}||{delay(1ms);exit(T); }
}
\end{verbatim}
    \end{scriptsize}

@@

\begin{scriptsize}

\begin{verbatim}
{
delay (1ms); emit S; //do something
}
\end{verbatim}
\end{scriptsize}
A periodic reaction (or CD) requires special consideration. Since the
delay statement is converted into \texttt{pause} constructs. One should
not introduce extra \texttt{pause} constructs when building periodic
reactions (or CDs). This is essential since, introducing pauses would
introduce more logical ticks.

@@

Preemption plays an important role in reactive languages. One needs to
carefully consider the interplay of \texttt{delay} semantics with the
preemption semantics of reactive languages. Other attempts at
incorporating delays (using external timers) have only been partially
successful, because of the complex interplay between real-time and
preemption. Consider the simple example below, which models real-time
using external timers as in~\cite{rsh94}.

\begin{scriptsize}

\begin{verbatim}
suspend{S} {
 emit START_TIMER(10); await (TIMER); 
 emit O1;
}
\end{verbatim}
\end{scriptsize}

As identified in~\cite{Bourke2009a} the \texttt{suspend} does not play
well with the external timer. The above program sends a signal to an
external timer and waits for 10ms to pass by. Consider what happens when
signals \texttt{S} and \texttt{TIMER} occur together, the \texttt{await}
statement is never executed and hence, we enter a deadlock. Such
problems are completely avoided in our technique, because we convert the
real-time delays into logical delays (\texttt{pause} constructs), which
bode well with preemption.

@@

Channels are an addition in the SystemJ language. Like interaction of
preemption and delays, conversion of rela-time delays to logical ticks
also bodes well with channel rendezvous, because the semantics of
interaction are well defined~\cite{amal10}. More importantly, we need to
consider the interplay of channel communication and WCRT/BCRT
analysis. Since, channel communication does not stop logical time
(see~\cite{amal10}) WCRT/BCRT are unaffected by channel
communication. The response time to input signals \textit{is} though!
But, we are unconcerned with the response time, analysis in this paper
and it remains a future research avenue.

>>>>

\section{Introducing real-time in SystemJ}
\label{sec:intr-real-time}

We introduce a single \textit{derived} (consisting of kernel constructs)
statement called \mbox{\texttt{delay (M..N)}} in the SystemJ language
for real-time control. The resolution of the delay statement is of
secondary concern and is dependent upon the execution platform.

@@

Given a SystemJ program: \texttt{delay (M..N); p}, where $M \in
\mathbb{Q}^{>0}$ and $N \in \mathbb{Q}^{>0}$, statement $p$ is executed
after real-time delay of $\tau$ units, such that, $M \leq \tau \leq N$.

We also introduce two variants of the derived statement \texttt{delay}.
\begin{enumerate*}
\item Given a SystemJ program \texttt{delay (M); p}, where $M \in
  \mathbb{Q}^{>0}$, statement $p$ is executed after real-time delay of
  $\tau$ units, such that, $M \leq \tau < \infty$. It is important to
  note that the lower bound of the delay construct $M$ is \textit{not}
  an exact delay, but rather the control is allowed to proceed to the
  next statement anytime after the delay of time $M$.
\item Given a SystemJ program \texttt{delay (M..M); p}, where $M \in
  \mathbb{Q}^{>0}$, statement $p$ is executed after real-time delay of
  $\tau$ units, such that, $M \leq \tau \leq M$. In this variant, the
  delay is exact.
\end{enumerate*}

The aforementioned variants are special cases of the general case
\texttt{delay (M..N)}.

@@

\begin{figure}[tb]
    \begin{minipage}{\textwidth}
      \begin{scriptsize}
\begin{verbatim}
trap(T){
 int x = 0;
 while(true){
  x = x + 1;
  pause;
  if(x == d) exit (T); //wait for "d" ticks
 } // d is the number of logical ticks calculated
} // by Algorithm 1
\end{verbatim}
      \end{scriptsize}
    \end{minipage}
    \caption{The rewrite of \texttt{delay} construct}
    \label{fig:3}
\end{figure}

@@

The computation of \texttt{d} is shown in Algorithm~\ref{alg:1}. This
algorithm is carried out for each CD (in SystemJ) or a synchronous
program individually. \textit{The fundamental observation is -- the
  reaction time for each logical tick is elastic -- varying only between
  the BCRT and the WCRT, thus any number of logical ticks \texttt{d}
  that map to the required real-time \texttt{delay} should be chosen in
  such a way that they are invariant to this elasticity.}

Our algorithm takes as input: WCRT, BCRT, and the lower and upper bounds
$M$ and $N$, respectively of the \texttt{delay} construct. We first
divide $M$ and $N$ with the BCRT and WCRT, respectively. This division
gives us the number of individual logical ticks required to delay the CD
(or synchronous program) by the real-time specification. We always
$ceil$ when dividing $M$ and $floor$ when dividing $N$ to make sure that
the resultant values are integers (in domain $\mathbb{N}^{>0}$) and
these functions guarantee that the resultant logical ticks result in
real-time delays between the required range $M-N$. Next, a function $F$
maps these calculated values to a set of equidistant integer points
(values) separated by a unit value -- these points represent all the
logical ticks running at the WCRT and the BCRT, respectively that
satisfy the real-time delay requirements. The intersection of these two
sets gives all the logical ticks that satisfy the real-time requirements
invariant of the reaction time and its elasticity.

Let us revisit our motivating example to illustrate the algorithm. From
Figure~\ref{fig:1} we know that $M$ is 50.3ms and $N$ is 200.3ms,
respectively. Let the WCRT and BCRT be: 0.112ms and 0.0334ms,
respectively. Thus, the algorithm proceeds as follows:

\begin{enumerate*}
\item $l_1 \leftarrow \lceil 50.3/0.112 \rceil$ and $u_1 \leftarrow
  \lfloor 200.3/0.112 \rfloor$. $l_1 = 450$ and $u_1 = 1788$. We first
  calculate the logical ticks that are always running at the WCRT
  satisfy the required real-time delay.
\item $l_2 \leftarrow \lceil 50.3/0.0334 \rceil$ and $u_2 \leftarrow
  \lfloor 200.3/0.0334 \rfloor$. $l_1 = 1506$ and $u_1 = 5997$. We do
  the same for the BCRT case.
\item $S_1 = [450..1788]$ and $S_2 =[1506,5997]$. We then map the
  resultant bounds to linear points. Sets $S_1$ and $S_2$ represent
  logical ticks that running at the WCRT and BCRT, respectively always
  satisfy the required real-time constraints.
\item $D = S_1 \cap S_2$, $D = [1506,1788]$. Finally, the intersection
  of the two sets gives the set $D$ from which we can pick up any value
  for \texttt{d}.
\end{enumerate*}

The resultant value for \texttt{d} gives the number of logical ticks,
which can run at any physical clock-speed, bounded by the BCRT and the
WCRT of the program and still result in the desired real-time delay. We
think this is an elegant solution, because the technique provides
hard-real time guarantee while preserving the essence of synchronous and
GALS programming prescribed by SystemJ and Esterel style
languages. Moreover, this technique considers non maximal-parallelism,
i.e., the delays in logical ticks is calculated after scheduling of CDs
or synchronous programs has been performed. To our knowledge we are the
first to do so.

@@

Many real-world systems require non-deterministic timing constructs. One
of them is our motivating example -- the human response time
system.

% Another is a printer-spooler example borrowed from timed
% CSP~\cite{Schneider:1999:CRT:555233}. The spooler and printer need to
% synchronize using channels. The printer might be unable to print
% depending upon the paper tray, similarly, spooler might take sometime
% to send the job depending upon its size. Such real-time constraints
% can be modeled in SystemJ as below:

% \begin{scriptsize}

% \begin{verbatim}
% {await (job); delay(2..10 ms);  //SPOOLER CD
%  send mid(job) delay (1ms);}
% ><
% {receive mid; delay(1..30 ms); //PRINTER CD
%  emit print(#mid);}
% \end{verbatim}
% \end{scriptsize}

@@

There are many instances when one would like to wait on an input from
environment for only a specified amount of time. This can be programmed
as:
% \begin{scriptsize}
\begin{verbatim}
// timeout after 1ms.
trap(T){{await(A);}||{delay((1..1)ms);exit(T);}}
\end{verbatim}
% \end{scriptsize}

@@

% \begin{scriptsize}
\begin{verbatim}
while(true) { delay ((1..1)ms); emit S; 
 //do something }
\end{verbatim}
% \end{scriptsize}
A periodic reaction (or CD) requires special consideration. Since the
delay statement is converted into \texttt{pause} constructs. One should
not introduce extra \texttt{pause} constructs when building periodic
reactions (or CDs). This is essential since introducing additional
pauses would introduce more logical ticks.

@@

Preemption plays an important role in reactive languages. One needs to
carefully consider the interplay of \texttt{delay} semantics with the
preemption semantics of reactive languages. Previous attempts at
incorporating delays (using external timers) have only been partially
successful, because of the complex interplay between real-time and
preemption. Consider the simple example below, which models real-time
using external timers as in~\cite{rsh94}.

% \begin{scriptsize}
\begin{verbatim}
suspend{S} {emit START_TIMER(10); await (TIMER); 
 emit O1;}
\end{verbatim}
% \end{scriptsize}

As identified in~\cite{Bourke2009a} the \texttt{suspend} does not play
well with the external timer. The above program sends a signal to an
external timer and waits for 10ms to pass by. Consider what happens when
signals \texttt{S} and \texttt{TIMER} occur in the same logical tick,
the \texttt{await} statement is never executed and hence, we enter a
deadlock. Such problems are completely avoided in our technique, because
we convert the real-time delays into logical delays (\texttt{pause}
constructs), which interact well with preemption.

@@

Channels, used for communication between reactions in asynchronously
running CDs, are an addition in the SystemJ language. Like interaction
of preemption and delays, conversion of rela-time delays to logical
ticks also interacts well with channel rendezvous, because the semantics
of interaction are well defined~\cite{amal10}. More importantly, we need
to consider the interplay of channel communication and WCRT/BCRT
analysis. Since, channel communication does not stop logical time
(see~\cite{amal10}) WCRT/BCRT are unaffected by channel
communication. % The response time to input signals \textit{is} though!
% But, we are unconcerned with the response time, analysis in this paper
% and it remains a future research avenue.
